---
title: "homework1.rmd"
author: "Sam Budoff"
date: "09/18/2014"
output: pdf_document
---
#Homework 1

## Cancer Data Frame

1. Load the data set into R and make it a data frame called `cancer.df`. (2 points)
```{r}
cancer.csv<-read.csv('cancer.csv')   # load cancer.csv
cancer.df <- data.frame(cancer.csv)  # convert into dataframe
```

2. Determine the number of rows and columns in the data frame. (2)
```{r}
nrow(cancer.df)  # count number of rows
ncol(cancer.df)  # count number of columns
```

3. Extract the names of the columns in `cancer.df`. (2)
```{r}
cnames <- colnames(cancer.df)  # create an array of character strings
cnames                         # display column names
```

4. Report the value of the 3000th row in column 6. (2)
```{r}
cancer.df[300,6]
```

5. Report the contents of the 172nd row. (2)
```{r}
cancer.df[172,]
```

6. Create a new column that is the incidence *rate* (per 100,000) for each row.(3)
```{r}
cancinc<-cancer.df$incidence                       # extract incidence
pop<-cancer.df$population                          # extract population
per<-(pop/10000)                                   # make population per 10000 individuals
cancrate<-cancinc/per                              #calculate rate per 10000
cancer2.df<-cbind(cancer.df, list(rate=cancrate))  # Add rate column, while renaming cancer.df
```

7. How many subgroups (rows) have a zero incidence rate? (2)
```{r}
sum(cancer2.df$rate ==0)
```

8. Find the subgroup with the highest incidence rate.(3)
```{r}
maxrate<- max(cancer2.df$rate)
rmaxrow <- function() {
  for (i in seq(1,42120,1)){   
    if (cancer2.df[i,9] == maxrate){
      mrownum<-print(i)
      }
    }
  }                                 # this for loop allows identification of the row number of interest
rmaxrow()                           # this command displays the row number of the subgroup of interst
```
**Running the function outputs the value 5797, which can be used to directly call the row of the dataframe with the highest incident rate**
```{r}
cancer2.df[5797,]
```

##Data Types

1. Create the following vector: `x <- c("5","12","7")`. Which of the following commands will produce an error message? For each command, Either explain why they should be errors, or explain the non-erroneous result. (4 points)

            max(x) 
*This command will produce an erroneous result as it will find the maximum value string of characters, in this case "7"*
            sort(x)
*This command will sort, again, on the basis of the character string value, not the implied numerical value, in this case* 
*"12", "5", "7"*
            sum(x)
*This command will result in an error message as characters can not be added*

2. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            y <- c("5",7,12)
*This concatonation will create a vector of character strings becasue the "5" vector is a character type and the default of*
*R is to make all concatonated arguments the same class type, with character having a higher priority than numberic*
            y[2] + y[3]
*This equation will fail because the numeric 7 and 12 are converted to character strings when concatonated with "5" in y.*
*Character strings can not undergo addition*

3. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            z <- data.frame(z1="5",z2=7,z3=12)
*This command results in teh creation of a 3 column data frame, where the first column is a list of character strings, and*
*the other columns are lists of numerics.*
            z[1,2] + z[1,3]
*This command will result in addition of the numerics 7 and 12 to yield the numeric 19. THis occurs becasue 7 and 12 are*
*the respective first rows of the second and third columns of the data frame z.*
                
##Data Structures
1. $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
```{r}
c(seq(1,8),seq(7,1))
```
    
2. $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$
```{r}
x=seq(1,5) 
rep(x,x)
```

3. $\begin{pmatrix}
      0 & 1 & 1  \\
      1 & 0 & 1  \\
      1 & 1 & 0  \\
    \end{pmatrix}$
```{r}
diag(-1,3,3)+matrix(1,3,3)
```

4. $\begin{pmatrix}
      1 & 2 & 3 & 4 \\
      1 & 4 & 9 & 16 \\
      1 & 8 & 27 & 64  \\
      1 & 16 & 81 & 256 \\
      1 & 32 & 243 & 1024  \\
    \end{pmatrix}$
```{r}
s=seq(1:4)
t(matrix(c(s,s^2,s^3,s^4,s^5),4))
```
##Basic Programming
1. Let $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$. Write an R program to calculate $h(x,n)$ using a `for` loop. (5 points)
**To use the following program type `Ztrans(x,n)` where x and n correspond, respectively, to x and n in $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$.**
```{r}
w=0
z=0
Ztrans<-function(x,n){
  for (i in seq(1,n)) {
  w=x^i
  z=z+w
  }
   print(z)
  }
```

2. If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6 and 9. The sum of these multiples is 23. Write an R program to perform the following calculations. (5 points)
```{r}
diviscounter<-function(x,y,z) {
  w=0 
  k=seq(1:(x-1))
   for (i in k) {
      if (i%%y==0) {
        w=i+w
      } else {  
      if (i%%z==0) {
        w=i+w
        }    
        }    
          }
      print(w)
      }
```

    a. Find the sum of all the multiples of 3 or 5 below 1,000. (3, [euler1])
*Using my function:*
```{r}
diviscounter(1e3,3,5)
```
    b. Find the sum of all the multiples of 4 or 7 below 1,000,000. (2)
*Using my function:*
```{r}
diviscounter(1e6,4,7)
```

3. Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first 10 terms will be $(1, 2, 3, 5, 8, 13, 21, 34, 55, 89)$. Write an R program to calculate the sum of the first 15 even-valued terms. (5 bonus points, [euler2])
**To operate this function subtract 2 from the Fibronaci sequence of interest's length, becasue 1 and 2 are implied in this function. For example, if you would like the sum of the the first 15 even terms wright* `fib(13)`**.**
```{r}
g=0
u=1
i=2
fib<- function(x) {
  while (x>0) {
    u=u+i  
    x=x-1
    if (u%%2==0) {g=g+u}
    i=i+u
    x=x-1
    if (i%%2==0) {g=g+i}
  }
  print(g)
}
fib(13)                     # print sum of first 15 even values in the Fibbonacci sequence
```